\documentclass[authoryear,preprint]{elsarticle}

\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{hyperref}
\usepackage{multirow}

\usepackage{defs}

\title{A Scalable Wave Resource Assessment Methodology: Application to U.S. Waters}
\author{Levi Kilcher, Gabriel Garc\'ia-Medina, Zhaoqing Yang}
\date{October 2019}

\begin{document}

\begin{abstract}
Wave energy technology research and development has accelerated over the last two decades. This is largely based on the recognition that waves deliver massive quantities of energy to populated coastlines around the world. Throughout this time, however, there have been ongoing debates about the correct methodology for quantifying the wave energy opportunity. The debate has centered on basic principals of wave energy converter technology, and the nature of the resource. In particular, the question of how to account for the regeneration of waves -- by wind -- down-wave from an array of wave energy converters, has not previously been addressed in detail. This has led to confusion and skepticism regrading the accuracy of existing methods, which has led to a split in wave energy resource assessment methodology depending on the user and scale of the assessment. Project developers seeking site assessments typically follow International Electrotechnical Commission technical specifications to quantify the opportunity at specific locations. On the other hand, regional resource assessments -- designed to quantify the bulk opportunity over large spatial scales -- have been conducted using a wide range of disparate methods. This work seeks to resolve the outstanding questions and debates in this field, and in the process develop a robust methodology that applies at both site-specific and regional scales. Applying this revised methodology to U.S. waters, we find the U.S. wave energy resource to be 3300 – 4100 TWh/yr, with region totals of: 2000 – 2500 TWh/yr in Alaska, 510 – 630 TWh/yr along the U.S. west coast, 380 – 470 TWh/yr along the east coast, 70 TWh/yr in the Gulf of Mexico, and 18-33 TWh/yr in Puerto Rico and the U.S. Virgin Islands.
\end{abstract}

\maketitle

\include{intro}
%\include{background}
\include{method}
\include{results}
\include{conclusion}

%\appendix

%\include{appendix-one-way}
%\include{appendix-flux-vs-area}

\clearpage

\bibliographystyle{elsarticle-num-names}
\bibliography{refs,refs_GGM}


\end{document}
